---
title: "`coretta2018itapol` data processing"
author: "Stefano Coretta"
date: "19/12/2017"
output: 
  html_document: 
    highlight: tango
    number_sections: yes
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(speakr)
library(beepr)
```

# Tangle and wave code

```{r tangle}
lmt("./data-raw/code/process.praat.md")
```

```{r wave}
system("pandoc -f gfm -o ./data-raw/code/process.pdf ./data-raw/code/process.praat.md -N --syntax-definition=/Users/stefano/.pandoc/pandoc-praat/praat.xml --highlight-style=/Users/stefano/.pandoc/flamenco.theme --metadata-file=./data-raw/code/process.yaml")
```

# Ultrasound data

## Force-alignment

**Export** `.wav` and `.txt` from `AAA` to `data-raw/data/ultrasound/derived/[ID]/recordings/` folder ([ID] = participant ID) as `[ID]-nnn.wav`.

Run `alignment-input.praat`. The output is a concatenated `.wav` file and a `.txt` file listing the stimuli, written in `data-raw/data/ultrasound/derived/[ID]/concatenated/`. Make sure the `.txt` contains 'Mowie' instead of 'Mówię' (unicode safe because of AAA).

Run SPPAS on the concatenated `.wav` and `.txt` files. **WARNING**: This procedure overwrites the files.

Perform manual correction of the alignment in `[ID]-palign.TextGrid` and save the corrected file to `data-raw/data/ultrasound/raw/corrected-palign/`.

Run `search-area.praat`. The output is a TextGrid file for each `.wav` file, saved in `data-raw/data/ultrasound/derived/[ID]/recordings/`, and a merged TextGrid, saved in `data-raw/data/ultrasound/derived/[ID]/concatenated/`.
**NOTE**: These files are needed for processing of the EGG data.

```{r search-area}
speakers <- c(
  "it01", "it02", "it03", "it07", "it09", "it11", "it12", "it13", "it14",
  "pl02", "pl03", "pl04", "pl05", "pl06", "pl07"
)
languages <- c(
  "it", "it", "it", "it", "it", "it", "it", "it","it",
  "pl", "pl", "pl", "pl", "pl", "pl"
)

for (i in 1:length(speakers)) {
  praat_run(
    "./data-raw/code/search-area.praat",
    "data-raw",
    speakers[i],
    languages[i]
  )
}
```

Run `closure-annotation.praat`. It reads the search area chunks and adds a tier for the time of closure onset.

```{r closure-annotation}
for (i in 1:length(speakers)) {
  praat_run(
    "./data-raw/code/closure-annotation.praat",
    "data-raw",
    speakers[i],
    languages[i]
  )
}
```

Now we can import the TextGrids back in `AAA`.

## Tongue tracking and kinematics

**Tongue tracking**:
  1. Create a spline template for the speaker and save it (`.fst`).
  1. Batch process splines with template.
  1. Check splines.
  1. (Batch process snap-to-fit).

**Calculate kinematics**
  1. Check relevant fan lines for Tongue Dorsum, Tongue Tip, and Tongue Root.
  1. Load `AnaVal-[ID].avl` from `Edit values` and change fan lines numbers (Save it for the future).
  1. Calculate Maths for entire session.

**Find consonantal gestures**:
  1. Select TD stimuli.
  1. `Find...` > `Load` and `execute` `td_function.srh`.
  1. `Find...` > `Load` and `execute` `find_gons_td.srh`.
  1. Select TT stimuli.
  1. `Find...` > `Load` and `execute` `tt_function.srh`.
  1. `Find...` > `Load` and `execute` `find_gons_tt.srh`.

**Export data**:
  1. `Export-tongue-cart.xsu`: export cartesian coordinates, tongue tip/dorsum/root smoothed displacement, velocity and absolute velocity at peaks, nucleus onset and offset, gesture onset, closure onset, and maximum (Annotation: "_"), `[ID]-tongue-cart.tsv` in `data-raw/datasets/ultrasound/`.
  1. `Export-vowel-series.xsu`: export cartesian coordinates, tongue tip/dorsum/root smoothed displacement, velocity and absolute velocity for every frame within the vowel (Annotation: "\^[aoOu]$"), `[ID]-vowel-series.tsv` in `data-raw/datasets/ultrasound/`.
  1. Export kinematics annotations (from `Export > Files...`) in `data-raw/data/derived/ultrasound/[ID]/kinematics` folder as `[ID].TextGrid`.

# Audio data

## C2 release detection

Run `burst-detection.praat`. The output is a `.TextGrid` file, written in `data-raw/data/ultrasound/derived/[ID]/concatenated/`. **WARNING:** This procedure overwrites corrected files!

Check detection with `[ID]-burst.TextGrid` and `[ID].wav` and overwrite `[ID]-burst.TextGrid`.

## C1 release detection

Run `release-detection-c1.praat`: the output is a `.TextGrid` file, written in `data-raw/data/ultrasound/derived/[ID]/concatenated/`.

```{r release-c1}
speakers <- c(
  "it01", "it02", "it03", "it04", "it05", "it07", "it09", "it11", "it12", "it13", "it14",
  "pl02", "pl03", "pl04", "pl05", "pl06", "pl07"
)
languages <- c(
  "it", "it", "it", "it", "it", "it", "it", "it", "it", "it","it",
  "pl", "pl", "pl", "pl", "pl", "pl"
)

for (i in 1:length(speakers)) {
    praat_run(
        "./data-raw/code/release-detection-c1.praat",
        speakers[i],
        languages[i]
    )
}
```

Check detection with `[ID]-realease-c1.TextGrid` and `[ID].wav`. Save corrected file in `data-raw/data/ultrasound/raw/corrected-palign/`.

## Get durations and formants

Run `get-durations.praat`. The output is written in `data-raw/datasets/acoustics/`.

```{r get-durations}
speakers <- c(
  "it01", "it02", "it03", "it04", "it05", "it07", "it09", "it11", "it12", "it13", "it14",
  "pl02", "pl03", "pl04", "pl05", "pl06", "pl07"
)
languages <- c(
  "it", "it", "it", "it", "it", "it", "it", "it","it", "it", "it",
  "pl", "pl", "pl", "pl", "pl", "pl"
)

for (i in 1:length(speakers)) {
    praat_run(
        "./data-raw/code/get-durations.praat",
        "data-raw",
        speakers[i],
        languages[i]
    )
}
```

Run `get-formants.praat`. The output is written in `data-raw/datasets/acoustics/`.

```{r get-formants}
sex <- c(
  "m", "m", "f", "f", "f", "m", "f", "m", "m", "f", "m",
  "f", "m", "f", "m", "m", "f"
)

for (i in 1:length(speakers)) {
    praat_run(
        "./data-raw/code/get-formants.praat",
        speakers[i],
        sex[i]
    )
}
```

# EGG data

Synchronise EGG files with `sync-egg.praat`. The output is written in `data-raw/data/egg/derived/[ID]/`.

```{r sync-egg}
speakers <- c(
  "it01", "it02", "it03", "it07", "it09", "it11", "it12", "it13", "it14",
  "pl02", "pl03", "pl04", "pl05", "pl06", "pl07"
)

for (i in 1:length(speakers)) {
    praat_run(
        "./data-raw/code/sync-egg.praat",
        "data-raw",
        speakers[i],
        0
    )
}
```

Extract vuv (Voiced/UnVoiced intervals) with `extract-vuv.praat`. The output is written in `data-raw/data/egg/derived/[ID]/`.

```{r vuv}
for (i in 1:length(speakers)) {
    praat_run(
        "./data-raw/code/extract-vuv.praat",
        "data-raw",
        speakers[i],
        40,
        10000,
        11,
        0
    )
}
```

Calculate dEGG tracegrams** with `degg-tracing.praat`. The results are written in `data-raw/data/datasets/egg/`.

```{r tracegram}
for (i in 1:length(speakers)) {
    praat_run(
        "./data-raw/code/degg-tracing.praat",
        "data-raw",
        speakers[i],
        40,
        10000,
        11
    )
}
```

Extract wavegram data with `wavegram.praat`.

```{r wavegram}
for (i in 1:length(speakers)) {
    praat_run(
        "./data-raw/code/wavegram.praat",
        speakers[i]
    )
}

beep()
```

Get voicing durations from EGG annotations with `voicing-duration.praat`.

```{r voicing-duration}
speakers <- c(
  "it01", "it02", "it03", "it07", "it09", "it11", "it12", "it13", "it14",
  "pl02", "pl03", "pl04", "pl05", "pl06", "pl07"
)

for (i in 1:length(speakers)) {
  praat_run("./data-raw/code/voicing-duration.praat", speakers[i])
}
```

# `AAA` files extensions

* `.esu`: tabs in the main window
* `.fst`: fans template
* `.avl`: analysis values chart
* `.srh`: search analysis value
* `.xsu`: export settings
